---
title: "AG3C_PCA"
output: html_document
---
## Abstract

## Results

### All Samples, All Conditions, All Time Points
```{r, fig.width=10, fig.height=5, echo=FALSE, message=FALSE}
##########################################################################
# This script must be run with 'deseq_normalized_mrna_data.csv' and
# 'metaRNA.csv' in the current working directory.
##########################################################################

rm(list=ls())
library(dplyr)
library(tidyr)
library(ggplot2)
library(gridExtra)

plot_pca <- function(filtered_data, color_by) {
  filtered_data %>% select(starts_with('ECB')) -> joined

  # Calculate PCA
  joined %>% prcomp() -> pca
  
  # Now you need to add back in a column with some information that will
  # be used to color the points on the PCA plot. Here I'm coloring by
  # growth conditions.
  pca_data <- data.frame(pca$x, condition=color_by)
  
  # Generate plot
  p1 <- ggplot(pca_data, aes(x=PC1, y=PC2, color=factor(condition))) + 
    geom_point() +
    coord_equal()
  
  # Let's look at our rotations
  rotation_data <- data.frame(pca$rotation, variable=row.names(pca$rotation))
  
  # define a pleasing arrow style
  arrow_style <- arrow(length = unit(0.05, "inches"),
                       type = "closed")
  
  # now plot, using geom_segment() for arrows and geom_text for labels
  p2 <- ggplot(rotation_data) + 
          geom_segment(aes(xend=PC1, yend=PC2), x=0, y=0, arrow=arrow_style) + 
          geom_text(aes(x=PC1, y=PC2, label=variable), hjust=0, size=3, color='red') + 
          xlim(-1.,1.25) + 
          ylim(-1.,1.) +
          coord_fixed()
  
  grid.arrange(p1,p2,ncol=2)
  
}

# The following few lines just import data from CSVs and merge it into one
# data frame for easy analysis.
reads <- read.csv('deseq_normalized_mrna_data.csv')
reads %>% gather(sample, count, -X) -> reads
metadata <- read.csv('metaRNA.csv')
joined <- inner_join(reads,metadata,by = c('sample'='dataSet')) %>%
  rename(gene=X) %>%
  spread(gene,count)

# Here is where you can filter the data based on various conditions
joined %>% filter() -> filtered
plot_pca(filtered, filtered$GrowthConditions)


```

This plot doesn't show much separation between the conditions, probably because there is such a diverse mixture of time points and growth conditions. However, the rotations show that ECB_00933, ECB_01646, ECB_02486, and ECB_03406 contribute the most to PC1 and PC2.

### All Samples, All Conditions, <24 Hours
```{r, fig.width=10, fig.height=5, echo=FALSE}
joined %>% filter(Growthtime.hr. < 24) -> filtered
plot_pca(filtered, filtered$GrowthConditions)


```

### All Samples, All Condtions, for each time point. 
```{r, fig.width=10, fig.height=5, echo=FALSE}
t_hr <- unique(joined$Growthtime.hr.)
for (t in t_hr)
{
  joined %>% filter(Growthtime.hr.==t) -> filtered
  if (length(unique(filtered$GrowthConditions))>1) plot_pca(filtered, filtered$GrowthConditions)
} 
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

## Methods

### Data Normalization

### Principal Component Analysis

We did data manipulations, PCA calculations, and data visualizations in R 3.1.2 [1]. We used two R packages, tidyr and dplyr, to format and filter data for given growth conditions and time points [2,3]. After filtering, we used prcomp() (R built-in function) with default arguments to do PCA. prcomp() uses singular value decomposition method, which is a preferred method for genome-wide expression data analysis [4]. For all PCA plots we used R package ggplot2 [5]. Our code for this section is available [here](https://github.com/dariyasydykova/AG3C_pca_mrna).     

### Gene Network Analysis


#### References

1. R. Ihaka, R Gentleman (1996). R: A Language for Data Analysis and Graphics. Journal of Computational and Graphical Statistics. 5:299-314
2. tidyr ???
3. H. Wickam (2009). plyr: Tools for splitting, applying and combining data. 
4. O. Alter, P. O. Brown and D. Botstein (2000). Singular Value Decomposition for Genome-Wide Expression Data Processing and Modeling. PNAS. 97:10101-10106
5. H. Wickam (2011). ggplot2. WIREs Comp Stat, 3: 180–185. doi: 10.1002/wics.147